---
title: "4_RDataTypes_2"
author: "sanudelhi1199"
date: "5/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Types

## Data Frame
### it stroe value in form of table and multiple class
#### excample

```{r Dataframe}

family = data.frame("name" = c("hitesh","gajuji","Manjulaben","Rashmika","Kamakhya"),"Relation" = c("Self","Father","Mother","Sister","Bhani"), "age" = c(22,50,48,30,3), "occupation" = c("Student","Farmer","Housewife","Housewife","NA"))

View(family)

print(family)

```

#### how to create data rame using vectors

```{r metrix}

name <- c("Hitesh","Gajuji")

age <- c(22,50)

occupation <- c("student", "Farmers")


data_frame_1 <- data.frame(name,age,occupation)


print(data_frame_1)



```

#### how to convert data frame to str

```{r strs}

str(data_frame_1)



```

#### how to enable disable factor 

```{r factor}

data_frame_1 <- data.frame(name,age,occupation, stringsAsFactors = T)

str(data_frame_1)

data_frame_1 <- data.frame(name,age,occupation, stringsAsFactors = F)

str(data_frame_1)

```

#### how to extract values form data frame 


```{r }

data_frame_1


data_frame_1[2,3]

data_frame_1$name[2]

data_frame_1$occupation[2]



```

#### how to get data from data from perticular

```{r}


data_frame_1[c(1,2),c("occupation","age")]


data_frame_2 <- mtcars


View(data_frame_2)


data_frame_2["Valiant",c("mpg","hp")]



```

#### diffrance between [] output in dta frame and [[]] output in vector


```{r}


data_frame_2[[1]] #### mpg in vector

data_frame_2[1] #### mpg in data frame


data_frame_2[["Valiant",c("mpg")]] #### vector formate

data_frame_2["Valiant",c("mpg")] #### data frame



```
#### how to add extra row and column in data frame


```{r}

name <- c("hitesh","gajuji")

age <- c(22,50)


dat_frame_3 <- data.frame(name,age)

dat_frame_3


occupation <- c("student","farmer")

dat_frame_3$occupation <- occupation

dat_frame_3

alive <- c(T,T)

cbind(dat_frame_3,alive)


df <- data.frame("name" = "rashmika", "age" = "30", "occupation" = "house-wife")

df


df2 <- rbind(dat_frame_3,df)

df2
```


#### how to short and order data

```{r}

df2

sort(df2$age)


ranks <- order(df2$age)

ranks

df2$age


ranks <- order(df$name)

ranks


```


#### how to order data frame

```{r}

df2

ranks <- order(df2$age)


df2[ranks,]



df2[order(df2$age, decreasing = F)]


```

# vectors 
## hold similar value one dimention

``` {r}

v1 <- c("Hitesh","gajuji","manjulaben")

v2 <- c(1,2,3)

v3 <- c(T,T,T)


v1;v2;v3

### extrenal coersion 


v4 <- c("hitesh",23,T)

v4


### creating sequence 


v5 <- 1:10
v5


v5 <- seq(1,10)
v5

v5 <- seq(1,10,5)
v5

v6 <- seq(from=21, by=5, length = 10)
v6


### how to assign value to vector


tempr <- c(45,44,43,42,41)

tempr

names(tempr) <- c("mon","tue","wed","thur","fri")

tempr

tempr["mon"]

tempr[1]


### ---------------------------------------------------------------

name <- c("hitesh","gajuji","rashmika","kamakhya")

age <- c(22,50,30,3)


names(age) <- name

age

### -------------------------------------------------------------


price <- c(100:110)

price


names(price) <- paste0("p",1:11)


price


### mathamatical operation on vector


v1 <- c(1,2,3)


v2 <- c(4,5,6)

v1 + v2


v1[1] + v2[3]


v1 * v2

v1 * v2[1]


### comparing the vectors



v1 <- c(4,5,6)

v2 <- c(7,8,9)

v1 > v2 

v2 > v1

v2[2] > v1[2]


v1 == v2

v1 != v2


v4 <- c(1,2,3,4,5,6,7,8)

v4 > 3

v4 != 3

v4 == 3


### --------------------------------------------


v5 <- c(101:110)

names(v5) <- paste0("p",1:10)

v5


v5[4] == v5["p4"]


### how to skeep value


v5[c(-2,-4)]


### ------------------------

v5

filter <- v5 > 104

v5[filter]

v5[v5>103]





```











