---
title: "4_RDataTypes_2"
author: "sanudelhi1199"
date: "5/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Types

## Data Frame
### it stroe value in form of table and multiple class
#### excample

```{r Dataframe}

family = data.frame("name" = c("hitesh","gajuji","Manjulaben","Rashmika","Kamakhya"),"Relation" = c("Self","Father","Mother","Sister","Bhani"), "age" = c(22,50,48,30,3), "occupation" = c("Student","Farmer","Housewife","Housewife","NA"))

View(family)

print(family)

```

#### how to create data rame using vectors

```{r metrix}

name <- c("Hitesh","Gajuji")

age <- c(22,50)

occupation <- c("student", "Farmers")


data_frame_1 <- data.frame(name,age,occupation)


print(data_frame_1)



```

#### how to convert data frame to str

```{r strs}

str(data_frame_1)



```

#### how to enable disable factor 

```{r factor}

data_frame_1 <- data.frame(name,age,occupation, stringsAsFactors = T)

str(data_frame_1)

data_frame_1 <- data.frame(name,age,occupation, stringsAsFactors = F)

str(data_frame_1)

```

#### how to extract values form data frame 


```{r }

data_frame_1


data_frame_1[2,3]

data_frame_1$name[2]

data_frame_1$occupation[2]



```

#### how to get data from data from perticular

```{r}


data_frame_1[c(1,2),c("occupation","age")]


data_frame_2 <- mtcars


View(data_frame_2)


data_frame_2["Valiant",c("mpg","hp")]



```

#### diffrance between [] output in dta frame and [[]] output in vector


```{r}


data_frame_2[[1]] #### mpg in vector

data_frame_2[1] #### mpg in data frame


data_frame_2[["Valiant",c("mpg")]] #### vector formate

data_frame_2["Valiant",c("mpg")] #### data frame



```
#### how to add extra row and column in data frame


```{r}

name <- c("hitesh","gajuji")

age <- c(22,50)


dat_frame_3 <- data.frame(name,age)

dat_frame_3


occupation <- c("student","farmer")

dat_frame_3$occupation <- occupation

dat_frame_3

alive <- c(T,T)

cbind(dat_frame_3,alive)


df <- data.frame("name" = "rashmika", "age" = "30", "occupation" = "house-wife")

df


df2 <- rbind(dat_frame_3,df)

df2
```


#### how to short and order data

```{r}

df2

sort(df2$age)


ranks <- order(df2$age)

ranks

df2$age


ranks <- order(df$name)

ranks


```


#### how to order data frame

```{r}

df2

ranks <- order(df2$age)


df2[ranks,]



df2[order(df2$age, decreasing = F)]


```













